---
title: "Activity 3 - MLR"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidymodels)
library(tidyverse)
library(GGally)
library(readxl)

```


```{r}

# Get sheet names
lake_sheets <- excel_sheets("~/STA631/Activities/activity03-mlr/data/buoy-2011-2020.xlsx")  

# Read all sheets to list
list_all <- lapply(lake_sheets, function(x) {       
  as.data.frame(read_excel("~/STA631/Activities/activity03-mlr/data/buoy-2011-2020.xlsx", 
                           sheet = x))})

# New names for sheets that follow R naming conventions
new_names <- c("meteor", "h20_tempf", "h20_tempc", "dissolved_02", 
               "depth", "chl_phyco", "notes")

# Rename list elements in list_all with new_names
names(list_all) <- new_names

# Extract list elements to Global Environment
list2env(list_all, envir = .GlobalEnv)

```




```{r}

# Change names

chl_phyco_names <- c("dates", "chlor", "phyco")
names(chl_phyco) <- chl_phyco_names

depth_names <- c("dates", "depth_2m", "depth_5m", "depth_8m", "depth_11m")
names(depth) <- depth_names

dissolved_02_names <- c("dates", "sat_2m", "conc_2m", "sat_5m", "conc_5m",
  "sat_8m", "conc_8m", "sat_11m", "conc_11m")
names(dissolved_02) <- dissolved_02_names

h20_tempf_names <- c("dates", "tnode_1.7m", "temp_2m", "tnode_4m", "temp_5m", 
  "tnode_6m", "tnode_8m", "tnode_10m", "temp_11m")
names(h20_tempf) <- h20_tempf_names

meteor_names <- c("dates", "air_temp_f", "humidity", "barom", "wind_speed",
  "max_wind_speed", "wind_direct", "rain_intensity", "tot_rain",  "annual_rain_accum", "inter_rain")
names(meteor) <- meteor_names

```


```{r}

lake_data <- dissolved_02 %>% inner_join(h20_tempf, by = "dates")

lake_data <- lake_data %>% inner_join(meteor, by = "dates")

```


1. Is this an observational study or an experiment?

**Answer:** This study is observational.

2. Describe the distribution of your response variable. Is the distribution skewed? Are there any other interesting/odd features (outliers, multiple peaks, etc.)?  Is this what you expected to see? Why, or why not?

**Answer:** The distribution of my response variable `concentration` is skewed to the left.
Noteworthy is that there appears to be a slight u-shape in that left tail. There is no temp reading at 8 meters in this dataset so I did not include the corresponding 8m concentration readings, so that could have something to do with it.

3. Excluding your response variable, select two other numeric variables (hint: look for <dbl> or <int> designations) and describe their relationship with each other using an appropriate visualization.

**Answer:** Are we choosing chlor or phyco? I have no idea what is going on in these scatterplots lol.

```{r}

lake_data <- lake_data %>% 
  pivot_longer(c(conc_2m, conc_5m, conc_11m), names_to = "conc_meters", 
               values_to = "concentration") %>% 
  pivot_longer(c(temp_2m, temp_5m, temp_11m), names_to = "temp_meters", 
               values_to = "temp_f") %>% 
  select(-c(conc_8m, sat_2m, sat_5m, sat_8m, sat_11m, tnode_1.7m, tnode_4m, tnode_6m, tnode_8m, tnode_10m))


```



```{r explore variables}

ggplot(lake_data) +
  geom_histogram(aes(x = concentration))

ggplot(lake_data) +
  geom_point(aes(x = temp_f, y = wind_speed))



```

```{r pairwise relationships}

lake_data %>% 
  select(concentration, temp_f, wind_speed) %>% 
  ggpairs()

```

4. For each pair of variables, how would you describe the relationship graphically? Do any of the relationships look linear? Are there any interesting/odd features (outliers, non-linear patterns, etc.)?

**Answer:** There appears to be a weak relationship between concentration and temperature. There may be a weak relationship between temperature and could be 

5. For each pair of variables, how would you describe the relationship numerically?

6. Are your two explanatory variables collinear (correlated)? Essentially, this means that adding more than one of these variables to the model would not add much value to the model. We will talk more on this issue in Activity 4 (other considerations in regression models).


```{r}

lm_spec <- linear_reg() %>%
set_mode("regression") %>%
set_engine("lm")

lm_spec

mlr_mod <- lm_spec %>% 
fit(concentration ~ temp_f + wind_speed, data = lake_data)

tidy(mlr_mod)



```

* Day 02

** Some Important Questions

1.  Is at least one of the $p$ predictors $X_1$, $X_2$, $\ldots$, $X_p$
    useful in predicting the response $Y$?
    
2.  Do all the predictors help to explain $Y$, or is only a subset of
    the predictors useful?
    
3.  How well does the model fit the data?

4.  Given a set of predictor values, what response value should we
    predict and how accurate is our prediction?
    

```{r}

# Drop conc_5m so qualitative variable conc_meters only has two levels
lake_data2 <- lake_data %>% 
  filter(conc_meters != 'conc_5m')

```



```{r}

# review any visual patterns
lake_data2 %>% 
  select(concentration, conc_meters, temp_f) %>% 
  ggpairs()

#fit the mlr model
lm_spec <- linear_reg() %>%
set_mode("regression") %>%
set_engine("lm")

mlr_mod <- lm_spec %>% 
fit(concentration ~ conc_meters + temp_f, data = lake_data2)

tidy(mlr_mod)


```

1. What is the label that R assigned to this explanatory variable term?

2. What information is represented here?

3. What information is missing here?

4. For each level of your qualitative variable, write the simplified equation of the estimated line for that level. Note that if your qualitative variable has two levels, you should have two simplified equations.

5. Interpret the parameter estimate for the reference level of your categorical variable in the context of your problem. Page 83 of the text can help here (or have me come chat with you).

6. Interpret the parameter estimate for your quantitative variable in the context of your problem.

7. Create a new model with the same response and quantitative explanatory variable, but now choose a qualitative variable with more than two (but, say, less than 5) levels and obtain the tidy model output. How does R appear to handle categorical variables with more than two levels?

```{r}


# review any visual patterns
lake_data %>% 
  select(concentration, conc_meters, temp_f) %>% 
  ggpairs()

#fit the mlr model
lm_spec2 <- linear_reg() %>%
set_mode("regression") %>%
set_engine("lm")

mlr_mod2 <- lm_spec2 %>% 
fit(concentration ~ conc_meters + temp_f, data = lake_data)

tidy(mlr_mod2)

```



